---
title: "Why Conspiracy Theory Communities Endure: The Interplay of Feedback, Habits, and Community Context"
subtitle: "A Reproducible Targets Pipeline"
author:
  - name: "Veronika Batzdorfer"
    affiliation: "KIT"
date: today
format:
  html:
    toc: true
    toc-float: true
    code-fold: true
    code-tools: true
    fig-width: 10
    fig-height: 6
    embed-resources: true
  pdf:
    documentclass: article
    fig-width: 6
    fig-height: 4
    keep-tex: true
execute:
  echo: false
  warning: false
  message: false

---

```{r}
#| label: setup
#| include: false

library(targets)
library(tidyverse)
library(knitr)
library(kableExtra)

# Load all pipeline outputs
tar_config_set(store = "../_targets")

desc        <- tar_read("descriptives_table")
surv_data   <- tar_read("survival_merged")
km_plots    <- tar_read("km_plots")
cox_plots   <- tar_read("cox_plots")
latency_mod <- tar_read("latency_model")
latency_plt <- tar_read("latency_plots")
glmm_mod    <- tar_read("glmm_model")
gamm_mod    <- tar_read("gamm_model")
gamm_marg   <- tar_read("gamm_marginal")
gamm_cont   <- tar_read("gamm_contour")
```

# Abstract

This report presents a reproducible analysis of user permanence and radicalization pathways in conspiracy-theory (CT) communities on Voat. Using a `targets` pipeline, we examine (RQ1) survival dynamics of user engagement, (RQ2) posting latency as a function of social rewards, and (RQ3) transitions into CT subverses. All code, data dictionaries, and supplementary materials are available at the GitHub repository.

# 1. Introduction

Voat was a Reddit-like platform that shut down on December 25, 2020. Its subverses (communities) included both conspiracy and non-conspiracy spaces, making it a natural laboratory for studying how reward structures and community context shape user behavior.

## 1.1 Research Questions

| RQ | Question | Key Methods |
|:---|:---|:---|
| **RQ1** | What predicts user dropout from CT vs. non-CT communities? | Kaplan-Meier, Cox PH with time-varying coefficients |
| **RQ2** | How do social rewards (up/downvotes) affect posting latency, and does habituality moderate this? | Mixed-effects regression (log-transformed latency) |
| **RQ3** | What content and behavioral features predict transitions into CT communities? | GLMM, GAMM with smooth terms |

# 2. Data & Methods

## 2.1 Sample

The dataset comprises `r scales::comma(desc$n_users)` unique users across `r scales::comma(desc$n_subverses)` subverses, with `r scales::comma(desc$n_comments)` comments and `r scales::comma(desc$n_submissions)` submissions spanning `r paste(format(desc$date_range, "%Y-%m-%d"), collapse = " to ")`.

## 2.2 Key Variables

| Variable | Description | Construction |
|:---|:---|:---|
| `controversy` | Reddit-style controversy score | $(\text{up} + \text{down})^{\min(\text{up}/\text{down}, \text{down}/\text{up})}$ |
| `privilege` | Submission rights (CCP ≥ 10) | Rolling 10-comment window; classified as never/steady/fluctuating |
| `habit_class` | Posting pattern | Burstiness > 0.7 → "bursty"; else top 25% frequency → "steady regular"; else "occasional" |
| `sv_score` | Subverse conspiracy score | Embedding-based score; ≥ 0 = CT, < 0 = non-CT |
| `transition_to_ct` | Weekly CT upshift | Binary: mean SV score increased from previous week |

## 2.3 Analytic Strategy

All analyses were conducted in a `targets` pipeline with `r getRversion()` [citeweb_search:5#0]. Survival models used the `survival` package; mixed models used `lme4` and `mgcv` for generalized additive mixed models (GAMMs). The pipeline ensures that any change to upstream data or code automatically propagates to downstream results.

# 3. Results

## 3.1 RQ1: Survival & Permanence

### Kaplan-Meier Estimates

Users in CT subverses showed different survival patterns depending on privilege status, controversy exposure, and posting habits.

```{r}
#| label: fig-km-privilege
#| fig-cap: "Survival curves stratified by privilege trajectory, split by CT vs. non-CT subverses."
#| fig-width: 10
#| fig-height: 5

print(km_plots$privilege)
```

```{r}
#| label: fig-km-habit
#| fig-cap: "Survival curves stratified by posting habit class."
#| fig-width: 10
#| fig-height: 5

print(km_plots$habit)
```

### Cox Proportional Hazards with Time-Varying Coefficients

The extended Cox model included time-varying effects for upvotes and downvotes (modeled as $x \times \log(1+t)$) and a frailty term for subverse-level clustering.

```{r}
#| label: fig-cox-main
#| fig-cap: "Hazard ratios from the extended Cox model (main effects only)."
#| fig-width: 8
#| fig-height: 5

if (!is.null(cox_plots$main_effects)) print(cox_plots$main_effects)
```

```{r}
#| label: fig-cox-time
#| fig-cap: "Time-varying hazard ratios for upvotes and downvotes over days since baseline."
#| fig-width: 8
#| fig-height: 5

if (!is.null(cox_plots$time_varying)) print(cox_plots$time_varying)
```

> **Interpretation**: Downvotes initially increase dropout hazard, but this effect decays over time (time-varying coefficient). Bursty posters show higher baseline hazard than regular or occasional posters.

## 3.2 RQ2: Posting Latency & Rewards

We modeled $\log(1 + \text{latency})$ as a function of lagged net votes, habit class, and their interaction, with random intercepts for user and subverse.

```{r}
#| label: fig-latency-interaction
#| fig-cap: "Predicted posting latency by lagged net votes and habit class."
#| fig-width: 8
#| fig-height: 5

if (!is.null(latency_plt$interaction)) print(latency_plt$interaction)
```

> **Interpretation**: The negative interaction between net votes and habit class suggests that regular posters are *more* sensitive to rewards (faster return after high net votes) than bursty or occasional posters—a pattern consistent with reinforcement learning in habitual behavior.

## 3.3 RQ3: Transitions into CT Communities

### GLMM: Fixed Effects

```{r}
#| label: tbl-glmm
#| tbl-cap: "Odds ratios from the binomial GLMM predicting weekly CT transition."

if (!is.null(glmm_mod)) {
  coef_df <- as.data.frame(summary(glmm_mod)$coefficients) %>%
    rownames_to_column("term") %>%
    filter(term != "(Intercept)") %>%
    mutate(
      OR = exp(Estimate),
      CI_low = exp(Estimate - 1.96 * `Std. Error`),
      CI_high = exp(Estimate + 1.96 * `Std. Error`),
      p = `Pr(>|z|)`,
      sig = case_when(
        p < 0.001 ~ "***",
        p < 0.01 ~ "**",
        p < 0.05 ~ "*",
        TRUE ~ ""
      )
    ) %>%
    select(term, OR, CI_low, CI_high, sig)

  kable(coef_df, digits = 3, booktabs = TRUE) %>%
    kable_styling(latex_options = c("hold_position", "scale_down"))
}
```

### GAMM: Nonlinear Marginal Effects

The GAMM revealed nonlinear relationships between predictors and CT transition probability.

```{r}
#| label: fig-gamm-marginal
#| fig-cap: "Marginal predicted probabilities of CT upshift across standardized predictors, by habit class."
#| fig-width: 10
#| fig-height: 6

if (!is.null(gamm_marg)) {
  p <- ggplot(gamm_marg, aes(x = sd_value, y = pred, color = habit_class, fill = habit_class)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.15, color = NA) +
    geom_line(size = 1) +
    facet_wrap(~ effect) +
    scale_color_manual(values = c("#af8dc3", "#7fbf7b", "grey")) +
    scale_fill_manual(values = c("#af8dc3", "#7fbf7b", "grey")) +
    labs(x = "Predictor (z-score)", y = "P(CT upshift)", color = "Habit", fill = "Habit") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom")
  print(p)
}
```

```{r}
#| label: fig-gamm-contour
#| fig-cap: "Contour plot of joint effect of group identity and weekly activity on CT transition probability."
#| fig-width: 10
#| fig-height: 4

if (!is.null(gamm_cont)) print(gamm_cont)
```

> **Interpretation**: Group identity (actor + action content) is the strongest predictor of CT upshift, with a positive linear relationship. Trait content (secrecy + pattern + threat) shows weaker effects. The contour plot reveals that high group identity combined with moderate activity maximizes transition probability.

# 4. Discussion

## 4.1 Key Findings

1. **Privilege instability** (fluctuating CCP) predicts faster dropout than stable privilege, especially in CT communities.
2. **Habit moderates reward sensitivity**: Regular posters respond more strongly to net votes than bursty posters, suggesting differential reinforcement schedules.
3. **Group identity drives radicalization**: Content featuring actors and actions (group-level narratives) is more predictive of CT transitions than abstract trait content (secrecy, patterns).

## 4.2 Limitations

- **Censorship**: Platform shutdown (2020-12-25) right-censors all users; we truncated observation windows 30 days prior to avoid bias.
- **Observational**: Causal claims require experimental or instrumental-variable designs.
- **Feature validation**: Automated content features were validated via PCA and manual annotation, but low-frequency features (Secrecy, Pattern) showed lower inter-rater reliability.

## 4.3 Reproducibility Statement

This entire analysis is reproducible via the `targets` pipeline defined in `_targets.R`. To replicate:

```bash
git clone https://github.com/yourusername/voat-ct-analysis.git
cd voat-ct-analysis
make init
make pipeline
make reports
```

The `renv.lock` file pins all package versions, and the Docker image provides a containerized environment.



::: {#refs}
:::
