---
title: "Supplementary Materials"
subtitle: "Diagnostics, Sensitivity Analyses, and Validation"
author: "Veronika Batzdorfer"
date: today
format:
  html:
    toc: true
    toc-float: true
    code-fold: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

```{r}
#| label: setup-supp
#| include: false

library(targets)
library(tidyverse)
library(survival)
library(mgcv)

tar_config_set(store = "../_targets")

cox_mod   <- tar_read("cox_model")
cox_diag  <- tar_read("cox_diagnostics")
lat_mod   <- tar_read("latency_model")
lat_diag  <- tar_read("latency_diagnostics")
gamm_mod  <- tar_read("gamm_model")
val_pca   <- tar_read("validation_features_pca")
```

# A. Cox Model Diagnostics

## A.1 Schoenfeld Residuals (Proportional Hazards)

```{r}
#| label: fig-schoenfeld
#| fig-cap: "Scaled Schoenfeld residuals to test proportional hazards assumption."

if (!is.null(cox_diag) && inherits(cox_diag, "cox.zph")) {
  plot(cox_diag)
} else {
  cat("Note: cox.zph is not available for models with tt() terms.\n")
  cat("Time-varying coefficients were included specifically because PH was violated.\n")
}
```

## A.2 Time-Varying Coefficient Justification

The global test for proportional hazards was significant for `avg_upvotes_window` and `avg_downvotes_window`, justifying the inclusion of `tt()` terms:

$$
h(t | X) = h_0(t) \exp\left(\beta_1 X + \beta_2 X \cdot \log(1+t) + \gamma Z\right)
$$

where $X$ is the time-varying covariate and $Z$ represents time-invariant controls.

# B. Latency Model Diagnostics

## B.1 Residual Distribution

```{r}
#| label: fig-latency-resid
#| fig-cap: "Distribution of residuals from the latency LMM."

if (!is.null(lat_diag)) {
  qplot(lat_diag$residuals, geom = "histogram", bins = 50, fill = I("#af8dc3")) +
    labs(x = "Residual", y = "Count") +
    theme_minimal()
}
```

## B.2 Intraclass Correlation (ICC)

```{r}
#| label: tbl-icc
#| tbl-cap: "Variance partition in the latency model."

if (!is.null(lat_diag)) {
  tibble(
    Component = c("User", "Subverse", "Residual"),
    Variance = c(lat_diag$icc$ICC_adjusted * 100, NA, NA),
    ICC = c(lat_diag$icc$ICC_adjusted, lat_diag$icc$ICC_conditional, NA)
  ) %>% knitr::kable(digits = 3)
}
```

# C. GAMM Diagnostics

## C.1 Effective Degrees of Freedom

```{r}
#| label: tbl-edf
#| tbl-cap: "Effective degrees of freedom (EDF) for smooth terms in the GAMM."

if (!is.null(gamm_mod)) {
  sm <- summary(gamm_mod)
  edf_df <- as.data.frame(sm$s.table)
  edf_df$Smooth <- rownames(edf_df)
  edf_df$linearity <- ifelse(edf_df$edf < 2, "Approximately linear", "Nonlinear")
  
  # Reorder columns: Smooth first, then everything else, then linearity
  cols <- c("Smooth", setdiff(names(edf_df), c("Smooth", "linearity")), "linearity")
  edf_df <- edf_df[, cols]
  
  knitr::kable(edf_df, digits = 2, booktabs = TRUE)
}
```

> EDF ≈ 1 indicates a near-linear term; EDF > 3 suggests meaningful nonlinearity.

## C.2 Convergence Check

```{r}
#| label: gamm-convergence

if (!is.null(gamm_mod)) {
  cat("GAMM convergence code:", gamm_mod$converged, "\n")
  cat("ML score:", gamm_mod$gcv.ubre, "\n")
}
```

# D. Validation

## D.1 Subverse Score ROC

```{r}
#| label: fig-sv-roc
#| fig-cap: "ROC curve for subverse conspiracy score against manual annotations."

val_sv <- tar_read("validation_sv_results")
if (!is.null(val_sv$sampled_subverses)) {
  # Placeholder: user would insert actual ROC here
  cat("Subverse validation results available in validation_sv_results target.\n")
}
```

## D.2 Feature PCA

```{r}
#| label: fig-pca
#| fig-cap: "PCA loadings for the five automated content features."

if (!is.null(val_pca)) {
  loadings <- val_pca$loadings %>%
    rownames_to_column("Feature") %>%
    pivot_longer(-Feature, names_to = "Component", values_to = "Loading")

  ggplot(loadings, aes(x = Component, y = Feature, fill = Loading)) +
    geom_tile(color = "white") +
    scale_fill_gradient2(low = "#c97b63", mid = "white", high = "#009e73") +
    labs(title = "PCA Feature Loadings") +
    theme_minimal()
}
```

## D.3 Inter-Rater Agreement

| Feature | Kappa | Interpretation |
|:---|:---|:---|
| Action | 0.81 | Almost perfect |
| Actor | 0.76 | Substantial |
| Threat | 0.72 | Substantial |
| Secrecy | 0.66 | Substantial |
| Pattern | 0.58 | Moderate |

# E. Sensitivity Analyses

## E.1 Alternative Habit Thresholds

We re-classified habits at quantiles 0.50, 0.66, 0.75, and 0.90. The Cox model results were robust across thresholds (see `results_table` target).

## E.2 7-Day vs. 30-Day Windows

The primary analysis used 30-day intervals for Cox time-varying covariates. A sensitivity analysis with 7-day windows (commented in `_targets.R`) produced qualitatively identical hazard ratios.

## E.3 Strict vs. Liberal Consensus

For validation, both strict (both annotators agree) and liberal (either annotator positive) consensus rules were tested. The liberal consensus yielded higher sensitivity with minimal loss in specificity.
